Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning

### 2018-02-07 paper.tex
progress on wake sleep

\newcommand{\system}{\textsc{DreamHacker}~}
\newcommand{\lowerBound}{\mathscr{L}}
% Use the following line for the initial blind version submitted for review:
\usepackage[nohyperref]{icml2018}

@@

\twocolumn[
\icmltitle{\system: Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}

@@

This work proposes \emph{learning} the DSL.
We consider the setting where we have a collection of related programming tasks,
each specified by a set of input/output examples.
We do not assume that the tasks are annotated with
ground-truth programs.
Our algorithm, called \system,
does two things:
it finds programs that solve each of the tasks,
and at the same time creates its own DSL by discovering and reusing
domain-specific subroutines.
The learned DSL
should distill the commonalities across the programs that solve the tasks,
and so can be used to solve other related program learning problems.

@@

How should we perform inference for the model
in~\ref{graphicalModel}(a)?  We take the intuitive strategy of exploiting the
conditional independence of the programs conditioned on the DSL, and
so alternate between estimating the DSL ($\mathcal{D},\theta$) and
estimating a posterior distribution over $p$ by searching for
programs.  In practice searching for
programs is difficult, so we use the recognition model to
learn how to search efficiently.

We apply \system to four domains:
 building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and Lisp-style functions on lists.
 For each of these we deliberately provide an impoverished
 set of programming primitives,
 and show that our algorithm discovers
 its own domain-specific vocabulary for expressing solutions in each of these domains (Table~\ref{initialExampleDSL}).
 \begin{table}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   Circuits&\texttt{(nand (nand a a) (nand b b))}\\
   &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \texttt{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \texttt{(join '' (split c s))}\\
   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
   Lists& \texttt{(map (lambda (x) (+ x a)) l)}\\
   &\hspace{0.5cm} \emph{(add a to every element of list l)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful fragments of code.}\label{initialExampleDSL}
 \end{table}

 \subsection{Related Work}
 Our work is far from the first to learn to learn programs,
 an  idea that goes back to Solomonoff~\cite{solomonoff1989system}:

@@

Why enumerate, when the program synthesis community has invented many
sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
We have two reasons:% for using the simple strategy of enumerating programs:
\begin{itemize}
\item A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
\item Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
  the space of of programs.
\end{itemize}
A main drawback of an enumerative search algorithm is that we have no
efficient means of solving for arbitrary constants that might occur in the
program. In Section~\ref{regressionSection},
we will show how to find programs with real-valued constants
by automatically differentiating through the program and setting the constants using gradient descent.
In Section~\ref{textSection}
we will show that the bottom-up neural recognition model can learn
which discrete constants should be included in a program.

@@

The purpose of the recognition model is to accelerate the search over
programs.  It does this by learning to predict which programs both 
assign high likelihood to a task, and at the same time 
have high prior probability under the prior $\probability[\cdot |\mathcal{D},\theta]$.

@@

\section{Model}

>>>>

\newcommand{\system}{\textsc{DreamCoder}~}
\newcommand{\lowerBound}{\mathscr{L}}
% Use the following line for the initial blind version submitted for review:
\usepackage[nohyperref]{icml2018}

@@

\twocolumn[
\icmltitle{Bootstrapping Domain Specific Languages for Neurally-Guided Bayesian Program Learning}

@@

This work proposes \emph{learning} the DSL.
We consider the setting where we have a collection of related programming tasks,
each specified by a set of input/output examples.
We do not assume that the tasks are annotated with
ground-truth programs.
Our algorithm, called \system,
does three things:
it finds programs that solve each of the tasks;
it creates its own DSL by discovering and reusing
domain-specific subroutines;
and it learns a neural network that
efficiently searches for programs in the DSL.
The purpose of the DSL
is to distill the commonalities across the programs that solve the tasks,
and it can be used to solve other related program learning problems.

@@

How should we perform inference for the model
in~\ref{graphicalModel}(a)?
We take inspiration from the Helmholtz machine and the Wake/Sleep algorithm~\cite{hinton1995wake},
which alternates between ``waking'', or training a generative model (our $\mathcal{D},\theta$),
and  ``sleeping'', or training a recognition model on samples from the generative model (our $q$).
For us, a wake cycle means using the recognition model to search for programs that solve the tasks,
while a sleep cycle means updating the DSL and the recognition model
to more closely match the distribution of tasks.
This wake/sleep style of inference is the reason why we call our approach \system.

We take the intuitive strategy of exploiting the
conditional independence of the programs conditioned on the DSL, and
so alternate between estimating the DSL ($\mathcal{D},\theta$) and
estimating a posterior distribution over $p$ by searching for
programs.  In practice searching for
programs is difficult, so we use the recognition model to
learn how to search efficiently.

We apply \system to four domains:
 building Boolean circuits; symbolic regression; FlashFill-style~\cite{gulwani2011automating} string processing problems; and Lisp-style functions on lists.
 For each of these we deliberately provide an impoverished
 set of programming primitives,
 and show that our algorithm discovers
 its own domain-specific vocabulary for expressing solutions in each of these domains (Table~\ref{initialExampleDSL}).
 \begin{table}
    \begin{tabular}{ll}
   \toprule
   Domain&Part of the learned DSL\\\midrule
   Circuits&\texttt{(nand (nand a a) (nand b b))}\\
   &\hspace{0.5cm} \emph{(an OR gate)}\\
   Regression& \texttt{(+ (* real x) real)} \\
      &\hspace{0.5cm} \emph{(a linear function of x)}\\
   Strings& \texttt{(join '' (split c s))}\\
   &\hspace{0.5cm} \emph{(delete occurrences of c in string s)}\\
   Lists& \texttt{(map (lambda (x) (+ x a)) l)}\\
   &\hspace{0.5cm} \emph{(add a to every element of list l)}
   \\\bottomrule
    \end{tabular}
    \caption{ Examples of structure found in DSLs  learned by our algorithm. \system builds a new DSL by discovering and reusing useful subroutines.}\label{initialExampleDSL}
 \end{table}

@@

Why enumerate, when the program synthesis community has invented many
sophisticated algorithms that search for programs?~\cite{solar2008program,schkufza2013stochastic,feser2015synthesizing,osera2015type,polozov2015flashmeta}.
We have two reasons:% for using the simple strategy of enumerating programs:
\begin{itemize}
\item A key point of our work is that learning the DSL, along with a neural recognition model, can make program induction tractable, even if the search algorithm is very simple.
\item Enumeration is a general approach that can be applied to any program induction problem. Many of these more sophisticated approaches require special conditions on
  the space of of programs.
\end{itemize}
A main drawback of an enumerative search algorithm is that we have no
efficient means of solving for arbitrary constants that might occur in the
program. In Section~\ref{regressionSection},
we will show how to find programs with real-valued constants
by automatically differentiating through the program and setting the constants using gradient descent.
%% In Section~\ref{textSection}
%% we will show that the bottom-up neural recognition model can learn
%% which discrete constants should be included in a program.

@@

The purpose of the recognition model is to accelerate the search over
programs.  It does this by learning to predict which programs both 
assign high likelihood to a task, and at the same time 
have high probability under $(\mathcal{D},\theta)$.

@@

\pagebreak
 \subsection{Related Work}
 Our work is far from the first to learn to learn programs,
 an  idea that goes back to Solomonoff~\cite{solomonoff1989system}:
